---
title: "Getting Data"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
```


The Stanford Open Policing Project data are available on a SQL server that you can query.  You may want to query a small piece and work with that file only.  However, you should practice using SQL to query the larger database. 


```{r}
SB <- readr::read_csv("ca_san_bernardino_2019_08_13.csv")
```

```{r}
SB %>% select(citation_issued, arrest_made) %>%
  table()
```




###  Connecting to the SQL server

```{r message = FALSE}
library(RMySQL)  
con <- dbConnect(
  MySQL(), host = "traffic.st47s.com", user = "student", 
  password = "Sagehen47", dbname = "traffic")
```



Next a series of SQL queries can be sent to the database using the `DBI::dbGetQuery()` function: each query returns an R dataframe.

```{r message = FALSE}
class(DBI::dbGetQuery(con, "SHOW TABLES"))
```

There are multiple tables within the `traffic` database.

```{r message = FALSE}
DBI::dbGetQuery(con, "SHOW TABLES")
```

The `EXPLAIN` command describes the ten field names (variables) in the `CAsb` table.

```{r message = FALSE, eval=FALSE}
DBI::dbGetQuery(con, "EXPLAIN CAsb")
```

The `SELECT` statement can be used to select all fields for eight observations in the `TNstate` table.

```{r message = FALSE}
eightobs <- DBI::dbGetQuery(con, "SELECT * FROM TNstate LIMIT 8")
eightobs
```


More interesting and complicated `SELECT` calls can be used to undertake grouping and aggregation.  Here we calculate the average age broken down by race.

```{r message = FALSE}
dbGetQuery(con, 
  "SELECT subject_race, AVG(subject_age) AS 'ave age' 
  FROM TNnashville 
  GROUP BY subject_race
  ORDER BY `ave age`")
```


#### Accessing a database using `dplyr` commands {-}

Alternatively, a connection can be made to the server by creating a series of `dplyr` tbl
objects. Connecting with familiar `dplyr` syntax is attractive because, as [Hadley Wickham](https://dbplyr.tidyverse.org/articles/dbplyr.html) has noted, SQL and R have similar syntax (but sufficiently different to be confusing).  

The setup process looks similar.

```{r message = FALSE}
db <- dplyr::src_mysql(dbname = "traffic", host = "traffic.st47s.com", user = "student", 
          password="Sagehen47")
CAsb <- dplyr::tbl(db, "CAsb")
TNstate <- dplyr::tbl(db, "TNstate")
```

We  explore the `CAsb` table using the `collect()` function used to force computation on the database (and return the results).  One attractive aspect of database systems is that they feature [*lazy evaluation*](https://en.wikipedia.org/wiki/Lazy_evaluation), where computation is optimized and postponed as long as possible.  

```{r message = FALSE}
CAsb  %>% dplyr::summarise(total = n())
CAsb %>%  dplyr::collect() %>% data.frame()   
# be careful with collect() when dealing with large tables!
```

Note how the number of rows is unknown (`10,000` at the top of the output above) for the lazy query.

Similarly, we can explore the `TNstate` table.

```{r message = FALSE}
TNstate  %>% dplyr::summarise(total = n())
TNstate %>% dplyr::collect()  # be careful with collect() with large tables!
```


