---
title: "sunrise and set"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(suncalc)
library(lubridate)
```

```{r}
CAsb <- read.csv("~/gitRepos/teaching/ResearchCircle2020/ca_san_bernardino_2019_08_13.csv")
str(CAsb)
```

1.  First I notice two things about the date/time.  (1) they are coded as factors (they should be dates, but at the very minimum, characters), (2) they are separate columns, and they might need to be combined.

NOTICE THE DIFFERENT TYPES OF VARIABLES IN THE COLUMN HEADERS!!!

```{r}
CAsb <- CAsb %>%
  dplyr::mutate(datetime = lubridate::ymd_hms(paste(as.character(date), as.character(time))),
                date2 = lubridate::as_date(lubridate::ymd(as.character(date)))) 

# check to make sure it did what we wanted it to do
CAsb %>%
  select(date, date2, time, datetime) %>%
  head()
```

2. Next I tried many of the bazillion different functions in R that will calculate the sunrise times.  They each had a very specific format for the date/time required.  Also, many of them had trouble with situations where lat and/or long was missing.  The function `getSunlightTimes` in the package `suncalc` seems to deal well with missing information.

For multiple lat/long/dates, the `data=` argument is required.

```{r}
temp <- data.frame(date = CAsb$date2, lat = CAsb$lat, lon = CAsb$lng)

getSunlightTimes(data = temp, keep = "sunrise") %>% 
  head()
```

3. Next I wrote a function (got this idea from the Stack Overflow page when looking at `sunriset`).

```{r}
oursunriseset <- function(latitude, longitude, date, direction = c("sunrise", "sunset")) {
  date.lat.long <- data.frame(date = date, lat = latitude, lon = longitude)
  if(direction == "sunrise"){
    getSunlightTimes(data = date.lat.long, keep=direction)$sunrise }else{
    getSunlightTimes(data = date.lat.long, keep=direction)$sunset  }
}
```

4. Let's check to make sure the function works how we think it should.  Remember that we need it to keep only the **one** column associated with the sunrise (or sunset).

```{r}
oursunriseset(CAsb$lat, CAsb$lng, CAsb$date2, direction = "sunrise") %>% head()
```


5. Let's use it in the mutate function!

```{r}
CAsb %>%
  dplyr::mutate(sunrise = oursunriseset(lat, lng, date2, direction = "sunrise"),
                sunset = oursunriseset(lat, lng, date2, direction = "sunset")) %>%
  select(date, time, date2, lat, lng, sunrise, sunset) %>%
  head()
```



